--- 
title: "Empowerment in Agriculture"
author: "Teresia Mrema Buza"
date: "Updated: `r Sys.time()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
always_allow_html: true
documentclass: book
bibliography: [book.bib, packages.bib, indexes.bib]
biblio-style: apalike
link-citations: yes
github-repo: tmbuza/indexbook
description: "Construction of empowerment indexes using the M0 model [@Alkire2011]."
show_downloads: true
favicon: "favicon.ico"
---

## Motivation
Strengthening reproducibility, portability and shareability. Here we will translate models developed elsewhere [@Alkire2011; @Alkire2012; @Alkire2013; @Malapit2019; ]. We are going to translate calculations for computing empowerment in agriculture indexes into R code. 

## Getting Started


## Methodology
```{block, type="tmbinfo", echo=TRUE}

1. Method 1 focuses on the percentage of empowered women and adequacies among the disempowered. 
2. Method 2 focuses on the percentage of disempowered women and the percentage of domains in which they lack adequate achievements. 

> Method 2 is commonly used as it is consistent with the M0 measurement (Alkire and Foster 2011).

* The disempowerment cut-off (k=20% ) is the percentage of weighted inadequacies a woman must have to be considered disempowered. 
* If ciw <= k the score is replaced by 0.
* Any existing inadequacies are not considered in the “Disempowered headcounts.
* Now the matrix has new column `ciwk` that contains censored inadequacy score.
```

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
